---
output: github_document
always_allow_html: yes
---
# cytoqc -- A standardization tool for openCyto

*cytoqc* checks and standardizes channels, markers, keywords, gates of the cytodata .


```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(knitr)
```

## Installation
```{r, eval=FALSE}
remotes::install_github("thebioengineer/colortable")
remotes::install_github("RGLab/cytoqc")
```

## Get started

```{r}
library(flowCore)
library(flowWorkspace)
library(cytoqc)
```

```{r include=FALSE}
# prepare the test data
data("GvHD")
fs <- GvHD[8:28]#exclude samples that has different Time length
data_dir <- tempfile()
dir.create(data_dir)
write.flowSet(fs, data_dir)
```

## Load the FCS

```{r}
files <- list.files(data_dir, ".fcs", full.names = TRUE)
cqc_data <- cqc_load_fcs(files)
cqc_data
```

```{r include=FALSE}
#simulate channel discrepancy

#case
cf <- cqc_data[[1]]
colnames(cf)[1] <- "fsc-h"

#missing
cqc_data[[1]] <- cf[,1:7]


#redundant
thisfile <- files[2]
fr <- read.FCS(thisfile)
new_col <- exprs(fr)[,8,drop=F]
colnames(new_col) <- "channelA"
fr <- fr_append_cols(fr, new_col)
write.FCS(fr, files[2])
cqc_data[[2]] <- load_cytoframe_from_fcs(thisfile, is_h5 = TRUE)

#typo
cf <- cqc_data[[2]]
colnames(cf)[2] <- "SSC1-H"

#both case and typo
cf <- cqc_data[[3]]
colnames(cf)[1:2] <- c("fsc-h", "SSC1-H")

#order
cf <- cqc_data[[4]]
cf_swap_colnames(cf, "FL1-H", "FL2-H")

```
## The basic workflow can be summarised as four steps:

1. check
2. match
3. fix

### 1. Check the consistency across samples
```{r}
check_results <- cqc_check(cqc_data, type = "channel")
check_results
```


### 2. Match the reference
```{r render=normal_print}
res <- cqc_match(check_results, ref = 3) 
res
```

### 3. Apply the fix

```{r}
cqc_fix(res)
```

## Update check report
```{r}
check_results <- cqc_check(cqc_data, type = "channel")
check_results
```

## Return the cleaned data
```{r}
cqc_data <- cqc_get_data(check_results)
cqc_data
```
## Coerce it inot `cytoset`
```{r}
cytoset(cqc_data)
```
## Or output to FCS
```{r eval=FALSE}
cqc_write_fcs(cqc_data, outdir)
```





