---
title: A QC tool for openCyto
output: rmarkdown::html_vignette
always_allow_html: yes
vignette: >
  %\VignetteIndexEntry{cytoqc-pregating}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# cytoqc -- A QC tool for openCyto

*cytoqc* performs pre-cleaning, pre-gating and post-gating QC checks.

## The basic workflow:

1. Run `cqc_group` to see the summary
2. `cqc_match_reference` and `cqc_find_solution` for selected groups
3. `cqc_fix` based on the solution or `cqc_drop_groups` for unsolvable issues
4. Iterate 1~3 steps until data is cleaned
5. `split` the data into multiple sets for multi-panel cases

```{r include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(kableExtra)
```


```{r}
library(flowCore)
library(flowWorkspace)
library(cytoqc)
```

```{r include=FALSE}
# prepare the test data
data("GvHD")
fs <- GvHD
data_dir <- tempfile()
dir.create(data_dir)
write.flowSet(fs, data_dir)
```

## Load the FCS

```{r}
files <- list.files(data_dir, ".fcs", full.names = TRUE)
cqc_data <- cqc_load_fcs(files)
cqc_data
```

```{r include=FALSE}
#simulate channel discrepancy

#case
cf <- cqc_data[[1]]
colnames(cf)[1] <- "fsc-h"

#missing
cqc_data[[1]] <- cf[,1:7]


#redundant
thisfile <- files[2]
fr <- read.FCS(thisfile)
new_col <- exprs(fr)[,8,drop=F]
colnames(new_col) <- "channelA"
fr <- fr_append_cols(fr, new_col)
write.FCS(fr, files[2])
cqc_data[[2]] <- load_cytoframe_from_fcs(thisfile, is_h5 = TRUE)

#typo
cf <- cqc_data[[2]]
colnames(cf)[2] <- "SSC1-H"

#both case and typo
cf <- cqc_data[[3]]
colnames(cf)[1:2] <- c("fsc-h", "SSC1-H")

#order
cf <- cqc_data[[4]]
cf_swap_colnames(cf, "FL1-H", "FL2-H")

```

## QC check for channel
```{r}
groups <- cqc_group(cqc_data, "channel")
su <- summary(groups)
su
diff(su)
```

## Match against the reference to find discrepancy


```{r}
match_result <-  cqc_match_reference(groups, ref = 3, select = c(1, 4))
match_result
```

## Recommend the solution  

```{r}
solution <- cqc_find_solution(match_result)
solution
```

*Show the itemized details*
```{r}
knit_print(solution, itemize = TRUE)
```

## Export/import the `solution` for revision (if needed)
```{r eval=FALSE}
library(readr)
write_csv(solution, csvfile)
#manually edit csvfile and load it back
solution_revised <- read_csv(csvfile)
```
## Apply the fix
```{r}
cqc_fix(solution)

```


## Refresh QC report 
```{r}
groups <- cqc_group(cqc_data, "channel")
summary(groups)
```

## Drop groups that can not be fixed
```{r}
groups <- cqc_drop_groups(groups, 1)
cqc_data <- cqc_get_data(groups)
length(cqc_data)
```

## QC for marker
```{r include=FALSE}
#simulate channel discrepancy

cf <- cqc_data[[1]]
cf_rename_marker(cf, "CD15 FITC", "cd15")
cf_rename_marker(cf, "CD33 APC", "apc cd33")


#redundant
cf <- cqc_data[[2]]
markernames(cf) <- c(`FL2-A` = "markerA")

```


```{r}
#summarise marker groups
groups <- cqc_group(cqc_data, "marker")
su <- summary(groups)
diff(su)
```

## Pick references and groups to fix
```{r}
solution <- cqc_find_solution(groups, ref = 3, select = 1, max.distance = 0.4) 
solution
cqc_fix(solution)

solution <- cqc_find_solution(groups, ref = 4, select = 2, max.distance = 0.4) 
solution
cqc_fix(solution)
```
## update checks
```{r}
groups <- cqc_group(cqc_data, "marker")
su <- summary(groups)
diff(su)
```

## split by groups
```{r}
data_list <- split(groups)
data_list
```

## QC for panel 
```{r}
# first data set is already clean
summary(cqc_group(data_list[[1]], "panel"))
```


## Save the cleaned data as FCS 
```{r}
cleaned <- data_list[[1]]
```


```{r eval=F}
out <- tempfile()
write_fcs(cleaned, out)
```


## Or Coerce it directly into `cytoset` (zero-copying)
```{r}
cs <- cytoset(cleaned)
cs
```

## Further split the data by panel
```{r}
grps <- cqc_group(data_list[[2]], "panel")
diff(summary(grps))
split(grps)
```

