---
title: cytoqc for lyoplate
output: rmarkdown::html_vignette
always_allow_html: yes
vignette: >
  %\VignetteIndexEntry{cytoqc-gs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# QC for tcell panel


```{r include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(kableExtra)
```


```{r}
library(flowCore)
library(flowWorkspace)
library(cytoqc)
```
```{r}
path <- "~/remote/fh/fast/gottardo_r/mike_working/lyoplate_out/parsed"
centers <- c('BIIR','CIMR','Miami','NHLBI','Stanford','UCLA','Yale')
```

## Load gs
```{r}
panel <- "tcell"
gslist <- sapply(centers, function(center) {
  message("Center: ", center)
  gs <- load_gs(file.path(path, center, panel))
  gs_clone(gs)
})

```

## QC Check gates
```{r}
cqc_data <- cqc_gs_list(gslist)

#group by gates
groups <- cqc_prepare(cqc_data, "gate")
su <- summary(groups)
knit_print(su)
knit_print(diff(su))

#vis the difference
plot_diff(groups)

# match reference
match_result <- cqc_match(groups, ref = 1, select = c(2))
match_result

solution <- cqc_recommend(match_result)
solution

cqc_fix(solution)

summary(cqc_prepare(cqc_data, "gate"))
```
## QC check for channel
```{r}
groups <- cqc_prepare(cqc_data, "channel")
su <- summary(groups)
knit_print(su)
diff(su)
```


## Channels are very different across centers so move on to check marker
```{r}
#summarise marker groups
groups <- cqc_prepare(cqc_data, "marker")
su <- summary(groups)
knit_print(su)
knit_print(diff(su))
```

## Markers are more standardized and go ahead to further clean it
```{r}
res <- cqc_match(groups, ref = 3, select = c(1,2,4,5,6)) 
res

solution <- cqc_recommend(res, max.distance = 0.6)
solution
cqc_fix(solution)

```
## update checks
```{r}
groups <- cqc_prepare(cqc_data, "marker")
knit_print(summary(groups))

```

## Relax string matching to clean up the rest
```{r}
res <- cqc_match(groups, ref = 2, select = c(1)) 
res

solution <- cqc_recommend(res, max.distance = 0.8)
solution
cqc_fix(solution)

groups <- cqc_prepare(cqc_data, "marker")
knit_print(summary(groups))
```

## Use the marker as reference to standardize the channels across centers
```{r}
panel <- cf_get_panel(gs_cyto_data(cqc_data[[1]])[[1]], skip_na = TRUE)
panel
cqc_set_panel(cqc_data, panel, ref.col = "marker")
groups <- cqc_prepare(cqc_data, "panel")
knit_print(summary(groups))
```

## Refresh QC report 
```{r}
groups <- cqc_prepare(cqc_data, "channel")
su <- summary(groups)
knit_print(su)
knit_print(diff(su))

```

## Remove `H/W` channels
```{r}
res <- cqc_match(groups, ref = 4, select = c(1,2,3)) 
res

solution <- cqc_recommend(res, max.distance = 0.6)
solution
cqc_fix(solution)
```

## Coerce it directly into single `GatingSet` (zero-copying)
```{r eval=FALSE}
gs <- merge_list_to_gs(cqc_data)
gs
```


