---
output: github_document
always_allow_html: yes
---
# QC for tcell panel


```{r include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(kableExtra)
```


```{r}
library(flowCore)
library(flowWorkspace)
library(cytoqc)
library(printr)
library(DT)
```
```{r}
path <- "~/remote/fh/fast/gottardo_r/mike_working/lyoplate_out/parsed"
centers <- c('BIIR','CIMR','Miami','NHLBI','Stanford','UCLA','Yale')
```

## Load gs
```{r}
panel <- "tcell"
gslist <- sapply(centers, function(center) {
  message("Center: ", center)
  gs <- load_gs(file.path(path, center, panel))
})

```

## QC Check gates
```{r}
cqc_data <- cqc_gs_list(gslist)

#group by gates
groups <- cqc_check(cqc_data, "gate")
groups
diff(groups)

#vis the difference
plot_diff(groups)

# match reference
match_result <- cqc_match(groups, ref = 1)
match_result

cqc_fix(match_result)

cqc_check(cqc_data, "gate")
```
## QC check for channel
```{r}
groups <- cqc_check(cqc_data, "channel")
groups
```


## Channels are very different across centers so move on to check marker
```{r}
groups <- cqc_check(cqc_data, "marker")
groups
```

## Markers are more standardized and go ahead to further clean it
```{r}
res <- cqc_match(groups, ref = 3) 
res
```

## Re-match by relaxing the string matching threshold
```{r}
res <- cqc_match(groups, ref = 3, max.distance = 0.6)
res
```

## Manually match the individual items that are still not matched
```{r}
res <- cqc_update_match(res, group = 1, map = c("CD4 PerCP-Cy55" = "CD4"
                                              , "CD8 APC-H7" = "CD8"
                                              , "Live Dead FITC" = "LIVE")
                        )
res
```


```{r}
cqc_fix(res)

```

## update checks
```{r}
groups <- cqc_check(cqc_data, "marker")
groups

```

## check pannel
```{r render=normal_print}
res <- cqc_check(cqc_data, "panel")
res
```


### Spread markers
```{r render=normal_print}
format(res, anchor = "marker")

```


## Use the marker as reference to standardize the channels across centers
```{r}
cf <- gs_cyto_data(cqc_data[[1]])[[1]]
panel <- cf_get_panel(cf, skip_na = TRUE)
panel
cqc_set_panel(cqc_data, panel, ref.col = "marker")
groups <- cqc_check(cqc_data, "panel")
groups
```

## Refresh QC report 
```{r}
groups <- cqc_check(cqc_data, "channel")
groups
diff(groups)

```

## Remove `H/W` channels
```{r}
res <- cqc_match(groups, ref = 4) 
res

cqc_fix(res)
```

## Coerce it directly into single `GatingSet` (zero-copying)
```{r}
gs <- merge_list_to_gs(cqc_data)
gs
```


